---
title: "Chapter 5"
author: "Mariana da Cruz Albertazzi"
date: "13/11/2020"
output: html_document
---
### Exercices 
#
#### 1. For the data InsectSpray, make a table for the number of insects for each spray with the mean, median and standard error.

```{r echo=FALSE} 
me <- med <- se <- In <- numeric()
sp <- character()

for(i in levels(InsectSprays$spray)){   #i is a random value given to the spray column.
  for(j in names(InsectSprays)[-2]){    #j is a random value given to the count column.
  
    sp <-c(sp,i)                   # vector that allows to go through the values related to the levels of spray.
   
    x <- InsectSprays[InsectSprays$spray ==i,j]   
    me <- c(me, mean(x))
    se <- c(se, sd(x) / sqrt(length(x)))
    med <- c(med, median(x))
    In <- c(In, sum(x))
    
  }}

d <- data.frame(Spray=sp, Insect_number=In, Media=me, Median=med, Standard_error=se)

#Making a table

knitr::kable(d <- data.frame (Spray=sp,
               Insect_number=In, 
               Media=me, 
               Median=med, 
               Standard_error=se),  
               caption = "Table 1. Effect of sprays on insect number.")

```
#
#### 2. Print a plot to see the differences of counts between sprays. Include a caption explaining the figure. Which type of plot is the one you choose and why?

```{r echo=FALSE, fig.cap = "Figure 1. Differences of counts between sprays."}
plot(count~spray, data = InsectSprays,
     ylab="Insect Number", xlab="Spray", 
     main = "Effect of sprays on insect number")
```  

#
#### 3. Test for differences between sprays using anova and a posthoc comparison and redo the previous plot including the representation of all posthoc differences.  

```{r echo=FALSE, fig.cap = "Figure 2. Statistically significant differences of counts between sprays using Tukey post-hoc test."}
an <- aov(count ~ spray, data = InsectSprays)
summary.aov(an)

tuk <- TukeyHSD(an)

agri <- agricolae::HSD.test(an, "spray", group = TRUE, console = TRUE)

plot(count~spray, data = InsectSprays,
     ylim= c(0,30),
     ylab="Insect Number", xlab="Spray", 
     main = "Effect of different sprays on the number of insects") 
     text(x=(1:6),y=28 , c("a","a","b","b","b","a"))

```

#
#### 4. Test for differences between sprays using non-parametric Kruskal-Wallis rank sum test. Again, redo the plot with these results. Note: Use agricolae::kruskal().

```{r echo=FALSE, fig.cap = "Figure 3. Statistically significant differences of counts between sprays using Kruskal Wallis test."}
 Krus <- agricolae::kruskal(InsectSprays$count, InsectSprays$spray)
 
     
     plot(count~spray, data = InsectSprays,
          ylim= c(0,30),
          ylab="Insect Number", xlab="Spray", 
          main = "Effect of different sprays on the number of insects") 
     text(x=(1:6),y=28 , c("a","a","c","b","bc","a"))
```
#
#### 5. Transform count data using sqrt(counts) and redo the anova, the Tukey posthoc comparison and the plot.

```{r echo=FALSE,fig.cap = "Figure 4. Statistically significant differences of counts between sprays using squareroot insect count and the Tukey post hoc test."}
sqrt(InsectSprays$count)
     
     an2 <- aov(sqrt(count) ~ spray, data = InsectSprays)
     summary.aov(an2)
     
     tuk2 <- TukeyHSD(an2)
     
     agri2 <- agricolae::HSD.test(an2, "spray", group = TRUE, console = TRUE)
     
     plot(count~spray, data = InsectSprays,
          ylim= c(0,30),
          ylab="Insect Number", xlab="Spray", 
          main = "Effect of different sprays on the number of insects") 
     text(x=(1:6),y=28 , c("a","a","c","b","bc","a"))
```
#
#### 6. Test for normality of residuals for the two performed anova analyses of points 4 and 6 using shapiro.test() and use plot the anova to see the qqplots and compare them.

```{r echo=FALSE, fig.cap = "Figure 5. Normality of residuals of the first Anova analyses using Shapiro test."}

shap <- shapiro.test(an[["residuals"]])
     shap2 <- shapiro.test(an2[["residuals"]])
     
     qqnorm(an[["residuals"]], main = "QQ plot for Anova 1 residuals")
     qqline(an[["residuals"]])

```
```{r echo=FALSE,fig.cap = "Figure 6. Normality of residuals of the second Anova analyses using Shapiro test."}
qqnorm(an2[["residuals"]], main = "QQ plot for Anova 2 residuals")
     qqline(an2[["residuals"]])
```
#
#### 7. Which of the previous analysis is the adequate in this case? Why? Is there any difference in the results between the square root transformed ANOVA and the Kruskal-Wallis analyses? Is there any difference in the results between the direct ANOVA and the square root transformed ANOVA? Which ones?

I consider that the Kruskal Wallis test is the most appropriate in this case because it groups the results in a more specific way than the Tukey post hoc test (which only divides the results into two groups: a and b). Furthermore, although both the Kruskal Wallis test and the Tukey post hoc test made after the squareroot insect count showed the same differences, I consider that it is preferable to use the first one because it uses a less complex code and because it does not modify the data of the count when raising them to the square root.  


